---
title: "Time Series Assignment"
author: "Sangamesh"

output: 
  pdf_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Question 1:
###Identify the ARIMA(p, d, q) model and the white noise variance estimate for the given data sets. (Use ts.plot, acf, pacf, eacf, arima, etc. Avoid using auto.arima except for verifying your answer.)

```{r}
# Read all data sets

TSD1 <- read.table("TSD1.txt", stringsAsFactors=FALSE)
TSD2 <- read.table("TSD2.txt", stringsAsFactors=FALSE)
TSD3 <- read.table("TSD3.txt", stringsAsFactors=FALSE)
TSD4 <- read.table("TSD4.txt", stringsAsFactors=FALSE)
TSD5 <- read.table("TSD5.txt", stringsAsFactors=FALSE)
TSD6 <- read.table("TSD6.txt", stringsAsFactors=FALSE)
TSD7 <- read.table("TSD7.txt", stringsAsFactors=FALSE)
TSD8 <- read.table("TSD8.txt", stringsAsFactors=FALSE)
TSD9 <- read.table("TSD9.txt", stringsAsFactors=FALSE)
TSD10 <- read.table("TSD10.txt", stringsAsFactors=FALSE)

summary(TSD1)
summary(TSD2)
summary(TSD3)
summary(TSD4)
summary(TSD5)
summary(TSD6)
summary(TSD7)
summary(TSD8)
summary(TSD9)
summary(TSD10)

as.numeric(TSD1$x)
as.numeric(TSD2$x)
as.numeric(TSD3$x)
as.numeric(TSD4$x)
as.numeric(TSD5$x)
as.numeric(TSD6$x)
as.numeric(TSD7$x)
as.numeric(TSD8$x)
as.numeric(TSD9$x)
as.numeric(TSD10$x)
```

```{r} 
#TSD1
mean1<-mean(TSD1$x)
sd1<-sd(TSD1$x)
ts.plot(TSD1)
acf(TSD1$x)
pacf(TSD1$x)
arima(TSD1$x)
```

```{r}
#TSD2
mean2<-mean(TSD2$x)
sd2<-sd(TSD2$x)
ts.plot(TSD2)
acf(TSD2)
pacf(TSD2)
arima(TSD2)
```

```{r}
#TSD3
mean3<-mean(TSD3$x)
sd3<-sd(TSD3$x)
ts.plot(TSD3)
acf(TSD3)
pacf(TSD3)
arima(TSD3)
```

```{r}
#TSD4
mean4<-mean(TSD4$x)
sd4<-sd(TSD4$x)
ts.plot(TSD4)
acf(TSD4)
pacf(TSD4)
arima(TSD4)
```

```{r}
#TSD5
mean5<-mean(TSD5$x)
sd5<-sd(TSD5$x)
ts.plot(TSD5)
acf(TSD5)
pacf(TSD5)
arima(TSD5)
```

```{r}
#TSD6
mean6<-mean(TSD6$x)
sd6<-sd(TSD6$x)
ts.plot(TSD6)
acf(TSD6)
pacf(TSD6)
arima(TSD6)
```

```{r}
#TSD7
mean7<-mean(TSD7$x)
sd7<-sd(TSD7$x)
ts.plot(TSD7)
acf(TSD7)
pacf(TSD7)
arima(TSD7)
```

```{r}
#TSD8
mean8<-mean(TSD8$x)
sd8<-sd(TSD8$x)
ts.plot(TSD8)
acf(TSD8)
pacf(TSD8)
arima(TSD8)
```

```{r}
#TSD9
mean9<-mean(TSD9$x)
sd9<-sd(TSD9$x)
ts.plot(TSD9)
acf(TSD9)
pacf(TSD9)
arima(TSD9)
```

```{r}
#TSD10
mean10<-mean(TSD10$x)
sd10<-sd(TSD10$x)
ts.plot(TSD10)
acf(TSD10)
pacf(TSD10)
arima(TSD10)
```



#Question 2:
###   Simulate a series of n = 500 Gaussian white noise observations as in and compute the sample ACF to lag 20. Compare the sample ACF you obtain to the actual ACF. Now repeat the same by using only n = 50. How does changing n affect the results?

```{r}
library(ggplot2)

#Population ACF

LAG = c(0:20) # lag up to 20
acf.p = c(1, rep(0,20)) # population acf (as above)
ACF.df.p = data.frame(lag = LAG, acf = acf.p) # turn the above into data frame

# n=500 gaussian white noise ACF:

w = rnorm(500,0,1) # n=500 gaussian white noise
ACF.500 = acf(w, lag.max = 20)
ACF.500 = acf(w, lag.max = 20, plot = FALSE)
ACF.df.500 = with(ACF.500, data.frame(lag, acf))

# n=50 gaussian white noise ACF:

w = rnorm(50,0,1) # n=50 gaussian white noise
ACF.50 = acf(w, lag.max = 20)
ACF.50 = acf(w, lag.max = 20, plot = FALSE)
ACF.df.50 = with(ACF.50, data.frame(lag, acf))

#combine the acfs (population, n=500, n=50)
dat = data.frame(rbind(ACF.df.500, ACF.df.50, ACF.df.p),
n = c(rep("n=500", 21), rep("n=50", 21), rep("population", 21)))

# comparsion:

ggplot(data = dat, mapping = aes(x = lag, y = acf)) +
geom_hline(aes(yintercept = 0)) +
facet_grid(n ~ . )+
geom_segment(mapping = aes(xend = lag, yend = 0))

```

###As we can see from above comnparison that as the sample size n gets larger the sample autocorrelation function gets close to its popoulation counterpart.


#Question 3:
###Consider the so2 data set, which is part of astsa package. Fit an ARIMA(p, d, q) model to the data, performing all of the necessary diagnostics. After deciding on an appropriate model, forecast the data into the future four time periods ahead (about one month) and calculate 95% prediction intervals for each of the four forecasts.

```{r}
library(astsa, quietly=TRUE, warn.conflicts=FALSE)
require(knitr)
library(ggplot2)
#Read in the data
so2

#Convert the data into a time series
so2 <- ts(so2)
so2

#plot the data
plot.ts(so2)

#decomposing non-seasonal data(smoothing)
library(TTR)
so2SMA <- SMA(so2, n=8) #moving average of n=8
plot.ts(so2SMA)

#Forecast:
so2F <-  HoltWinters(so2, gamma = F)
so2F
so2F; so2F$SSE
plot(so2F) #observed v/s fitted

#Forecast for 4 time periods
library("forecast")
so2F2 <- forecast:::forecast.HoltWinters(so2F, h=4)
so2F2
plot(forecast(so2F2))

acf(so2F2$residuals, lag.max=20, na.action = na.omit)
Box.test(so2F2$residuals, lag=20, type='Ljung-Box')
plot.ts(so2F2$residuals)
```

## ARIMA model :

```{r}
#Differencing a time series
#First differencing
so2Diff <- diff(so2, differences = 1)
plot.ts(so2Diff)
#The time series of first differences (above) does appear to be stationary 
#in mean and variance, as the level of the series stays roughly constant 
#over time, and the variance of the series appears roughly constant over time. 

#selecting a candidate ARIMA Model
# Install the tools
source(url("http://lib.stat.cmu.edu/general/tsa2/Rcode/itall.R"))

# Plot ACF and PACF
acf(so2Diff, lag.max=20)            # plot a correlogram
acf(so2Diff, lag.max=20, plot=FALSE) # get the autocorrelation values
pacf(so2Diff, lag.max=20)             # plot a partial correlogram
pacf(so2Diff, lag.max=20, plot=FALSE) # get the partial autocorrelation values

library(forecast)
auto.arima(so2)
```

## Forecasting Using an ARIMA Model:

```{r}

so2arima <- arima(so2, order=c(1,1,2)) # fit an ARIMA(1,1,2) model
so2arima

library("forecast") # load the "forecast" R library
so2forecasts <- forecast:::forecast.Arima(so2arima, h=4, level=c(95))
so2forecasts
plot(forecast(so2forecasts))

```


# Question 4:
### Perform a time series model specification, estimation, model diagnostics, and forecasting for a financial stock of your choice.
```{r echo=f, message=f, error=f, warning=FALSE}
library(quantmod);
library(tseries);
library(timeSeries);
library(forecast);
library(xts);


```

```{r echo=f, message=f, error=f, warning=FALSE}
# Pull data from Yahoo finance 
getSymbols('MSFT', from='2000-01-01', to='2017-11-01')

# Select the relevant close price series
stock_prices = MSFT[,4]


# Compute the log returns for the stock
stock = diff(log(stock_prices),lag=1)
stock = stock[!is.na(stock)]

# Plot log returns 
plot(stock,type='l', main='log returns plot')


# Conduct ADF test on log returns series
print(adf.test(stock))

# Split the dataset in two parts - training and testing
breakpoint = floor(nrow(stock)*(2.9/3))

# Apply the ACF and PACF functions
par(mfrow = c(1,1))
acf.stock = acf(stock[c(1:breakpoint),], main='ACF Plot', lag.max=100)
pacf.stock = pacf(stock[c(1:breakpoint),], main='PACF Plot', lag.max=100)


# Initialzing an xts object for Actual log returns
Actual_series = xts(0,as.Date("2017-11-01","%Y-%m-%d"))

# Initialzing a dataframe for the forecasted return series
forecasted_series = data.frame(Forecasted = numeric())

for (b in breakpoint:(nrow(stock)-1)) {
  
  stock_train = stock[1:b, ]
  stock_test = stock[(b+1):nrow(stock), ]
  
  # Summary of the ARIMA model using the determined (p,d,q) parameters
  fit = arima(stock_train, order = c(2, 0, 2),include.mean=FALSE)
  #summary(fit)
  
  # plotting a acf plot of the residuals
  #acf(fit$residuals,main="Residuals plot")
  
  # Forecasting the log returns
  arima.forecast = forecast:::forecast.Arima(fit, h = 1,level=99)
  #summary(arima.forecast)
  
  # plotting the forecast
  #par(mfrow=c(1,1))
  #plot(arima.forecast, main = "ARIMA Forecast")
  
  # Creating a series of forecasted returns for the forecasted period
  forecasted_series = rbind(forecasted_series,arima.forecast$mean[1])
  colnames(forecasted_series) = c("Forecasted")
  
  # Creating a series of actual returns for the forecasted period
  Actual_return = stock[(b+1),]
  Actual_series = c(Actual_series,xts(Actual_return))
  rm(Actual_return)
  
  #print(stock_prices[(b+1),])
  #print(stock_prices[(b+2),])
  
}

# Adjust the length of the Actual return series
Actual_series = Actual_series[-1]

# Create a time series object of the forecasted series
forecasted_series = xts(forecasted_series,index(Actual_series))
```

```{r echo=f, message=f, error=f, warning=FALSE}
# Create a plot of the two return series - Actual versus Forecasted
plot(Actual_series,xlim = c(1000,1000),type='l',main='Actual Returns Vs Forecasted Returns')
lines(forecasted_series,lwd=1.5,col='red')
legend('topleft',c("Actual","Forecasted"),lty=c(1,1),lwd=c(1.5,1.5),col=c('black','red'))

# Create a table for the accuracy of the forecast
comparsion = merge(Actual_series,forecasted_series)
comparsion$Accuracy = sign(comparsion$Actual_series)==sign(comparsion$Forecasted)
print(comparsion)

# Compute the accuracy percentage metric
Accuracy_percentage = sum(comparsion$Accuracy == 1)*100/length(comparsion$Accuracy)
print(Accuracy_percentage)
```


