---
title: "Assignment_C"
author: "Sangamesh"
date: "4 December 2018"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q.1] 1. Perform the following steps and comment on the observation.

Step I. Generate one U(-100,100) random number. Call it m
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
m <- floor(runif(1, -100, 100))
print(paste0("m = ", m))
```
\hfill\break
Step II. Generate one U(10,50) random number. Call it s.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
s <- floor(runif(1, 10, 50))
print(paste0("s = ", s))
```
\hfill\break
Step III. Generate one U(10,25) random number. Call it n.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
n <- floor(runif(1, 10, 25))
print(paste0("n = ", n))
```
\hfill\break

Step IV. Generate 1000 N(m,s) random numbers. Call this the population.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
population <- rnorm(1000, m, s)
print(head(population))
```
\hfill\break

Step V. Sample n numbers without replacement from the population.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
sample <- sample(population, n, replace = FALSE)
print(paste("Head of sample: ", sample))
```
\hfill\break

Step VI. Construct 90%, 95%, and 99% confidence intervals for the population mean.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(BSDA)
ninty <- z.test(population, sigma.x = s, conf.level = 0.90)
print("For 90% interval: ")
print(ninty)

nintyFive <- z.test(population, sigma.x = s, conf.level = 0.95)
print("For 95% interval: ")
print(nintyFive)

nintyNine <- z.test(population, sigma.x = s, conf.level = 0.99)
print("For 99% interval: ")
print(nintyNine)
```
\hfill\break

Step VII. Construct 90%, 95%, and 99% confidence intervals for the population variance.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(EnvStats)
df <- length(population) - 1
print(paste("degree of freedom = ", df))
varPopulation <- var(population)
print(paste0("Population variance = ", varPopulation))

print("For 90% interval: ")
print(varTest(population, conf.level = 0.9, sigma.squared = s))

print("For 95% interval: ")
# lowerNintyFive = varPopulation * df / qchisq(0.05/2, df, lower.tail = FALSE)
# upperNintyFive = varPopulation * df / qchisq(1 - 0.05/2, df, lower.tail = FALSE)
# print(c(lower = lowerNintyFive, variance = varPopulation, upper = upperNintyFive))
print(varTest(population, conf.level = 0.95, sigma.squared = s))

print("For 99% interval: ")
print(varTest(population, conf.level = 0.99, sigma.squared = s))
```
\hfill\break

Step VIII. Repeat steps V & VI 100/500/1000 times and count the number of times (and percentage) that the population mean is captured by the confidence interval.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
draws = matrix(rnorm(floor(n) * 100, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/100 * 100
paste("For n=100: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 500, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/500 * 100
paste("For n=500: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 1000, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/1000 * 100
paste("For n=1000: Count = ", count, ", Percentage = ", percentage, "%")
```
\hfill\break

Step IX. Repeat steps V & VII 100/500/1000 times and count the number of times (and percentage) that the population variance is captured by the confidence interval.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
draws = matrix(rnorm(floor(n) * 100, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/100 * 100
paste("For n=100: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 500, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/500 * 100
paste("For n=500: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 1000, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/1000 * 100
paste("For n=1000: Count = ", count, ", Percentage = ", percentage, "%")
```
\hfill\break

Q.2] In a filament cut test, a razor blade was tested six different times with ultimate forces corresponding to 8.5, 13.9, 7.4, 10.3, 15.7, 4.0.
 
a] find 95% confidence interval on mean using standard t-distribution
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
forces <- c(8.5, 13.9, 7.4, 10.3, 15.7, 4.0)
forcesNintyFive <- t.test(forces, sigma.x = sd(forces), conf.level = 0.95)
print("For 95% interval: ")
print(forcesNintyFive)
```
\hfill\break
b] Find a 95% confidence interval on the mean using Efron's percentile method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(boot)
fboot <- function(x, i) mean(x[i])
bs <- boot(forces, fboot, R=1000)
efronForcesPerc <- boot.ci(bs, conf=0.95, type="perc")
print(efronForcesPerc)
```
\hfill\break

c] Find a 95% confidence interval on the mean using the BCa method and the ABC method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(boot)
fbootBCa <- function(x, i) mean(x[i])
bsBCa <- boot(forces, fbootBCa, R=1000)
efronForcesBCa <- boot.ci(bsBCa, conf=0.95, type="bca")
print("BCa test")
print(efronForcesBCa)

print("ABC test")
fabc <- function(x, w) w%*%x
efronForcesABC <- abc.ci(forces, fabc, conf =0.95)
print(efronForcesABC)
```
\hfill\break

d] Find a 95% confidence interval on the mean using the percentile-t method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(resample)
bootC <- bootstrap(forces, mean, seed = 0)
percentileForces <- CI.percentile(bootC, probs = c(0.025, 0.975))
# percentileForces <- CI.percentile(bootC, probs = c(0.95))
print("percentile-t  test:")
print(percentileForces)
```
\hfill\break

```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
```

```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
```

```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
```

