---
title: "Assignment1"
author: "Sangamesh"
date: "3 September 2018"
output:
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q.1


Calculating the Sample mean and draw it's histogram:
```{r x ,echo=FALSE}
inv_cdf <- function(u){k <- (1/u)^(1/3)}
u <- runif(1000, min=0, max=1)
x <- inv_cdf(u)
#x
```

```{r SampleMeanX,echo=FALSE}
hist(x)
SampleMeanX <- mean(x)
```
Sample mean of X is:
```{r,echo=FALSE}
print(SampleMeanX)
```
Calculating Expected value of the distribution:
```{r expectedX,echo=FALSE}
f1 <- function(x)(3/x^3)
expectedX = integrate(f1, 1, Inf)
```
Expected value of X is:
```{r,echo=FALSE}
print(expectedX)
```
Difference between sample mean and expected value of X is:
```{r,echo=FALSE}
Difference <- SampleMeanX - 1.5
print(Difference)
```
Inference : The difference between the sample mean and expected value is nearly 0.

## Q.2


```{r,echo=FALSE}
xv <- (1:100)
fv <- ((pi^(xv/2))/ gamma((xv/2)+1))*1^xv
#print(fv)
plot(xv,fv,type='h')
```
Observation: if the limit as d goes to infinity, the volume of the ball goes to zero.

## Q.3 


```{r,echo=FALSE}
xsa <- (1:50)
fsa <- (2*(pi ^(xsa/2))/ gamma(xsa/2))
#print(fsa)
plot(xsa,fsa,type='h')
```
Observation: if the limit as d goes to infinity, the surface area of the ball goes to zero.

## Q.4

```{r,echo=FALSE}
r <- (1:10)
d <- (1:10)
f1 <- function(r,d)((pi^(d/2))/ gamma((d/2)+1))*r^d
z <- outer(r,d,f1)
persp(r,d,z,theta=30, phi=30, expand = 0.5, col="lightblue")
```
Observation: if the limit as d and r goes to z, the volume of the ball goes to zero.

## Q.5 

```{r,echo=FALSE}
rsa <- (1:10)
dsa <- (1:10)
fsa <- function(rsa, dsa){(2*(pi ^(dsa/2))/ gamma(dsa/2))*(rsa^dsa)}
zsa <- outer(rsa,dsa,fsa)
persp(rsa,dsa,zsa,theta=30, phi=30, expand = 0.5, col="lightblue")
```
Observation: if the limit as d and r goes to infinity, the surface area of the ball goes to zero.


## Q.6

```{r sphere, echo = FALSE}
library(hitandrun)

sample <- hypersphere.sample(900, 20)
fin_sample <- 30*sample
```

```{r distance, echo = FALSE}
d <- dist(fin_sample,method = "euclidean")
```

```{r function, echo = FALSE}
scaling <- function(d,ndim){
  fit <- cmdscale(d, eig = FALSE, k = ndim)
  d_new <- dist(fit)
  diff <- (((k^0.5)*d) - d_new)
  print(paste("Subspace of dimension:",k))
  print(max(diff))
}
```

The differences in euclidean distances for the subspace projection is as follows:
```{r 1dim,echo = FALSE}
k = 1
scaling(d,k)
```

```{r 2dim,echo = FALSE}
k = 2
scaling(d,k)
```

```{r 3dim,echo = FALSE}
k = 3
scaling(d,k)
```

```{r 4dim,echo = FALSE}
k = 4
scaling(d,k)
```

```{r 5dim,echo = FALSE}
k = 5
scaling(d,k)
```

```{r 10dim,echo = FALSE}
k = 10
scaling(d,k)
```

Observation : As we decrease the dimension, the differences in euclidean distances for the subspace projection decreases.


[Attached FIle for code](https://drive.google.com/file/d/1yB7lo8myCUKfkqiwbj0dqArPmtGiqCwt/view?usp=sharing)



