---
title: "Teleco"
author: "Sangamesh"
date: "24 September 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PHASE 1: Reading the Dataset
```{r input, echo=FALSE}

data <- read.csv("teleco.txt", header = TRUE, sep = "\t")
#print(tail(data))
```

# PHASE 2 => Data Pre-Processing

Step 1: Checking for any Missing Data and Statistical Summary
```{r missing, echo = FALSE, warning=FALSE, message=FALSE}
library(Hmisc)
library(dplyr)
library(ggpubr)
details <- describe(data) 
details
```

So, we find missing data in 4 columns namely logtoll, logequi, logcard, logwire.

Since all the data that have missing values are of [log of the user], we will replace the missing values with mean of the data.

```{r echo = F}
for(i in 1:ncol(data)){
  data[is.na(data[,i]), i] <- mean(data[,i], na.rm = TRUE)
}
print(describe(data))
```
Result: So no missing values present now.

Step 2: Checking for any outliers by histogram method.
```{r echo=F}
#for (col in 2:ncol(data)) {
 #   hist(data[,col], main=paste(colnames(data[col]), sep = ""))
#}
for (col in 2:(ncol(data)-1)) {
    if(max(data[col]) > 1){
    boxplot(data[,col], main=paste(colnames(data[col]), sep = ""))}
}
```

Result : There are outliers in many columns but it is not necessary that all the parameters of the same row are having extreme data, so we are not removing any data from it.

Step 3: Normality test
```{r echo=F, messages=F ,warning=F}
library(Hmisc)
library(dplyr)
library(ggpubr)
#shapiro.test(data$churn)
```
```{r echo=F,warning=F}

for (col in 2:(ncol(data)-1)) {
    if(max(data[col]) > 1){
    scaled_data <- scale(data[,col], center = TRUE, scale = TRUE)
    #gghistogram(scaled_data)
    hist(scaled_data, breaks =50, main=paste(colnames(data[col]), sep = ""))
    # Normality check
    #print(colnames(data[col]))
    #print(shapiro.test(scaled_data))
}
}
```

# EDA
```{r echo=F, messages=F, warning=F}
ggdensity()
```
# PCA 
```{r warnings=F,echo=F, messages=F}
library(psych)
pca <- principal(data, nfactors = 4, residuals = FALSE, rotate="none", scores=TRUE)
pca
#pca$values
#plot(pca$values, type = "b")
```

Result: As a result of PCA,
COmponents having highest variability are custcat, equipmon, equipten, marital