---
title: "Assign1_1"
author: "Sangamesh"
date: "4 September 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading of data

Data files with n1 = 10^6, 10^7 and 5x10^7 are generated and their files are also created. But for n1 = 10^8, the memory reached limit and the PC stopped responding.



```{r data1, echo=FALSE}
n1 <- 1000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)), file = "data14a.txt", sep = " ", quote=T)
```

```{r data2, echo=FALSE}
n1 <- 10000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)), file = "data14b.txt", sep = " ", quote=T)
```

```{r data3, echo=FALSE}
n1 <- 50000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)), file = "data14c.txt", sep = " ", quote=T)
```

<!--

```{r data4, echo=FALSE}
n1 <- 100000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon
data14d<-cbind(y,x1,x2,x3,x4,x5)
fwrite(as.data.frame(data14d), file = "data14d.txt", sep = " ", quote=T)
```
-->

```{r data,echo=FALSE}
library(data.table)
data1 = fread("data14a.txt")
data2 = fread("data14b.txt")
data3 = fread("data14c.txt")
#data4 = fread("data14d.txt")
```

#lm command

```{r lm1, echo=FALSE}
gc(reset = TRUE)
start.time<-proc.time()
lm1 = lm(y~., data1)
summary(lm1)
end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```

```{r lm2, echo=FALSE}
gc(reset = TRUE)
start.time<-proc.time()
summary(lm(y~., data2))
end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```

- lm command worked for n1 = 10^6 and 10^7 data files, but the memory allocation reached limit for further data files.

#biglm command

```{r biglm1, echo=FALSE}
gc(reset = TRUE)
start.time<-proc.time()
summary(biglm(y~x1+x2+x3+x4+x5, data1))
end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```

```{r biglm2, echo=FALSE}
gc(reset = TRUE)
start.time<-proc.time()
summary(biglm(y~x1+x2+x3+x4+x5, data2))
end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```


- biglm command worked for n1 = 10^6 and 10^7 data files, but the memory allocation reached limit for further data files. But in comparison with lm command, it is way faster in processing, as being shown by the RAM allocation and "Number of minutes running".


#BigMemory & BigAnalytics

```{r bigmemory, echo=FALSE}
gc(reset = TRUE)
start.time<-proc.time()

n1 <- 1000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon


write.big.matrix(as.big.matrix(cbind(y,x1,x2,x3,x4,x5)), file = "data14Tempa.txt", sep = " ")

end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running to write: ", save.time[3]/60, "\n\n")

start.time<-proc.time()

w<-read.big.matrix("data14Tempa.txt", header=F, sep=" ", col.names = c("y","x1","x2","x3","x4","x5"))

end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running to read: ", save.time[3]/60, "\n\n")

start.time<-proc.time()

summary(biglm.big.matrix(y~x1+x2+x3+x4+x5, w))

end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running to find biglm: ", save.time[3]/60, "\n\n")

gc()
```


```{r biganalytics1, echo=FALSE}

gc(reset = TRUE)
start.time<-proc.time()

w<-read.big.matrix("data14b.txt", header=F, sep=" ", col.names = c("y","x1","x2","x3","x4","x5"))

summary(bigglm.big.matrix(y~x1+x2+x3+x4+x5, w))

end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```


```{r biganalytics2, echo=FALSE}

gc(reset = TRUE)
start.time<-proc.time()

w<-read.big.matrix("data14c.txt", header=F, sep=" ", col.names = c("y","x1","x2","x3","x4","x5"))

summary(biglm.big.matrix(y~x1+x2+x3+x4+x5, w))

end.time<-proc.time()
save.time<-end.time-start.time
cat("\nNumber of minutes running: ", save.time[3]/60, "\n\n")
gc()
```


```{r netcdf1, echo=FALSE}
library(biglm)
library(ncdf4)

n1 <- 1000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon

ncpath <- "C:/Users/hp/Documents/Google Drive NIIT/Semester 7/Suman Sanyal/Predictive Modelling/Assign3/data/"
ncname <- "ncdTemp1"
ncfile <- paste(ncpath, ncname,".nc4",sep="")
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)),ncfile, row.names=FALSE, sep=" ")
```


```{r netcdf2, echo=FALSE}
library(biglm)
library(ncdf4)

n1 <- 10000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon

ncpath <- "C:/Users/hp/Documents/Google Drive NIIT/Semester 7/Suman Sanyal/Predictive Modelling/Assign3/data/"
ncname <- "ncdTemp2"
ncfile <- paste(ncpath, ncname,".nc",sep="")
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)),ncfile, row.names=FALSE, sep=" ")
```


```{r netcdf3, echo=FALSE}
library(biglm)
library(ncdf4)

n1 <- 50000000
x1 <- 1:n1
x2 <-runif(n1,5,95)
x3 <- rbinom(n1,1,.4)
x4 <- rnorm(n1, mean=-30, sd=200)
x5 <- runif(n1,-5000,5000)
b0 <- 17; b1 <- -0.466; b2 <- 0.037; b3 <- -5.2; b4 <- 2; b5 <- 0.00876
sigma <- 1.4
epsilon <- rnorm(x1,0,sigma)
y <- b0 + b1*x1  + b2*x2  + b3*x3 + b4*x4 + b5*x5 + x1*x2 + epsilon

ncpath <- "C:/Users/hp/Documents/Google Drive NIIT/Semester 7/Suman Sanyal/Predictive Modelling/Assign3/data/"
ncname <- "ncdTemp3"
ncfile <- paste(ncpath, ncname,".nc",sep="")
fwrite(data.frame(cbind(y,x1,x2,x3,x4,x5)),ncfile, row.names=FALSE, sep=" ")
```