---
title: |
  | Inferential Statistics for Data Science
  | DS412
  | Kotalwar Sangamesh
  | U101115FCS210
  | Course in-charge: Prof. Suman Sanyal
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
  html_document:
    df_print: paged
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center}\LARGE\includegraphics[width=12cm]{logo.png}\\[\bigskipamount]}
- \posttitle{\end{center}}
---
\newpage
\tableofcontents
\newpage

#Acknowledgement
I'm highly indebted to Prof. Suman Sanyal for his guidance and constant supervision as well as for providing necessary information regarding the assignments.
&nbsp;

I acknowledge that any work that I submit for assessment at NIIT University:
&nbsp;

1. Must be all my own work.
&nbsp;

2. Must not have been prepared with the assistance of any other person, except those permitted within University guidelines or the specific assessment guidelines for the piece of work.
&nbsp;

3. Has not previously been submitted for assessment at this University or elsewhere.


\newpage

&nbsp;


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r wrap-hook, echo =FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

#Assignment A

##Q.1] How close is the average of the samples to the expected value of X?


Calculating the Sample mean and draw it's histogram:
```{r x ,echo=FALSE}
inv_cdf <- function(u){k <- (1/u)^(1/3)}
u <- runif(1000, min=0, max=1)
x <- inv_cdf(u)
#x
```

```{r SampleMeanX,echo=FALSE}
hist(x)
SampleMeanX <- mean(x)
```
Sample mean of X is:
```{r,echo=FALSE}
print(SampleMeanX)
```
Calculating Expected value of the distribution:
```{r expectedX,echo=FALSE}
f1 <- function(x)(3/x^3)
expectedX = integrate(f1, 1, Inf)
```
Expected value of X is:
```{r,echo=FALSE}
print(expectedX)
```
Difference between sample mean and expected value of X is:
```{r,echo=FALSE}
Difference <- SampleMeanX - 1.5
print(Difference)
```
Inference : The difference between the sample mean and expected value is nearly 0.

##Q.2] 2-D The volume of a d-dimensional unit ball


```{r,echo=FALSE}
xv <- (1:100)
fv <- ((pi^(xv/2))/ gamma((xv/2)+1))*1^xv
#print(fv)
plot(xv,fv,type='h')
```
Observation: if the limit as d goes to infinity, the volume of the ball goes to zero.

##Q.3] 2-D The surface area of a d-dimensional unit ball


```{r,echo=FALSE}
xsa <- (1:50)
fsa <- (2*(pi ^(xsa/2))/ gamma(xsa/2))
#print(fsa)
plot(xsa,fsa,type='h')
```
Observation: if the limit as d goes to infinity, the surface area of the ball goes to zero.

##Q.4] 3-D The volume of a d-dimensional unit ball

```{r,echo=FALSE}
r <- (1:10)
d <- (1:10)
f1 <- function(r,d)((pi^(d/2))/ gamma((d/2)+1))*r^d
z <- outer(r,d,f1)
persp(r,d,z,theta=30, phi=30, expand = 0.5, col="lightblue")
```
Observation: if the limit as d and r goes to z, the volume of the ball goes to zero.

##Q.5] 3-D The surface area of a d-dimensional unit ball

```{r,echo=FALSE}
rsa <- (1:10)
dsa <- (1:10)
fsa <- function(rsa, dsa){(2*(pi ^(dsa/2))/ gamma(dsa/2))*(rsa^dsa)}
zsa <- outer(rsa,dsa,fsa)
persp(rsa,dsa,zsa,theta=30, phi=30, expand = 0.5, col="lightblue")
```
Observation: if the limit as d and r goes to infinity, the surface area of the ball goes to zero.


##Q.6] Calculate distance

```{r sphere, echo = FALSE}
library(hitandrun)

sample <- hypersphere.sample(900, 20)
fin_sample <- 30*sample
```

```{r distance, echo = FALSE}
d <- dist(fin_sample,method = "euclidean")
```

```{r function, echo = FALSE}
scaling <- function(d,ndim){
  fit <- cmdscale(d, eig = FALSE, k = ndim)
  d_new <- dist(fit)
  diff <- (((k^0.5)*d) - d_new)
  print(paste("Subspace of dimension:",k))
  print(max(diff))
}
```

The differences in euclidean distances for the subspace projection is as follows:
```{r 1dim,echo = FALSE}
k = 1
scaling(d,k)
```

```{r 2dim,echo = FALSE}
k = 2
scaling(d,k)
```

```{r 3dim,echo = FALSE}
k = 3
scaling(d,k)
```

```{r 4dim,echo = FALSE}
k = 4
scaling(d,k)
```

```{r 5dim,echo = FALSE}
k = 5
scaling(d,k)
```

```{r 10dim,echo = FALSE}
k = 10
scaling(d,k)
```

Observation : As we decrease the dimension, the differences in euclidean distances for the subspace projection decreases.


[Attached File for code](https://drive.google.com/file/d/1yB7lo8myCUKfkqiwbj0dqArPmtGiqCwt/view?usp=sharing)

\newpage

#Assignment B
**The app can be run on cloud [here!](https://sangamkotalwar.shinyapps.io/Assignmnet_2Final/)**
```{r echo=T, eval=FALSE}
library(shiny)
# Define UI for application that draws a histogram
ui <- fluidPage(
  tags$head(
    tags$style(HTML("
                    .mg-histogram .mg-bar rect {
                    fill: #75AADB;
                    shape-rendering: auto;
                    }
                    .mg-histogram .mg-bar  {
                    fill: #75AADB;
                    shape-rendering: auto;
                    }
                    
                    
                    .mg-histogram .mg-bar rect.active {
                    fill: #ffa500;
                    }"))),
  
  # Sidebar with a slider input for number of bins 
  sidebarLayout(
    sidebarPanel(   selectInput("typee","Select the type of distribution",c("Uniform Distribution"="unii","Normal Distribution"="normm","Exponential Distribution"="expp"),selected = "Unifrom Distribution"),
                    selectInput("sele","Replacement",c("Yes"="yess","No"="noo")),
                    
                    sliderInput("numb",
                                "Number of Samples:",
                                min = 20,
                                max = 100,
                                value = 20),
                    sliderInput("unimin",
                                "Enter the Minimum Value for for Uniform Distibution: ",
                                min = 1,
                                max = 99,
                                value = 1),
                    sliderInput("unimax",
                                "Enter the Maximum Value for Uniform Distibution: ",
                                min = 2,
                                max = 100,
                                value = 100),
                    sliderInput("meann",
                                "Enter the Mean for Normal Distibution: ",
                                min = 0,
                                max = 50,
                                value = 10),
                    sliderInput("stddev",
                                "Enter the Standard Deviation for Normal Distibution: ",
                                min = 0,
                                max = 50,
                                value = 10),
                    
                    sliderInput("exprate",
                                "Enter the Rate for Exponential Distibution: ",
                                min = 1,
                                max = 100,
                                value = 1),
                    tags$div(class="header", checked=NA,
                             tags$p("Check the deployed version here:"),
                             tags$a(href="https://sangamkotalwar.shinyapps.io/Assignmnet_2Final/", "Click Here!")
                    )
    ),
    
    
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("plot",brush = brushOpts(id = "plot_brush"),hover = hoverOpts(id = "plot_hover"))
    )
  )
    )

# Define server logic required to draw a histogram
server <- function(input, output) {
  d <- reactive({
    
    sele <- switch(input$sele,
                   noo = TRUE,
                   yess = FALSE,
                   TRUE)
    disttt<-switch(input$typee,
                   unii=sample(input$unimin:input$unimax,input$numb,replace = sele),
                   normm=rnorm(input$numb,mean=input$meann,sd=input$stddev),
                   expp=rexp(input$numb,input$exprate),
                   sample(input$unimin:input$unimax,input$numb,replace = sele)
    )

  })
  faa<- reactive({
    disttt<-switch(input$typee,
                   unii="Uniform Distribution of ",
                   normm="Normal Distribution of ",
                   expp="Exponential Distribution of ",
                   "Uniform Distribution of "
    )
  })
  dkk<-reactive({

    if( is.null(input$plot_brush$xmax) && is.null(input$plot_hover$x)) 
      color="blue"
    else if(!is.null(input$plot_hover$x))
    {
      color=dkkb2()
    }
    else if( !is.null(input$plot_brush$xmax) && is.null(input$plot_hover$x)) 
    {
      color=dkkb()
    }
    
    else color=dkkb()
  })
  
  dkkb<-reactive({
    color="blue"
    flag=1
    i=1
    differe =((max(d())-min(d()))/10)
    check=min(d())
    while(i<11)
    {  
      if (check>(input$plot_brush$xmax))
      {
        flag=2
      }
      if(((input$plot_brush$xmin-differe)<check) && (flag==1))
      {
        color[[i]]<-"orange"
      }  
      else{
        color[[i]]<-"blue"
      }
      i=i+1
      check=check+differe
    }
    check=min(d())
    
    
    return(color)
  })
  dkk2<-reactive({

    if( is.null(input$plot_hover$x) ) 
    {
      color="blue"
    }
    else
      color=dkkb2()
    
  })
  dkkb2<-reactive({
    color=c("blue","blue","blue","blue","blue","blue","blue","blue","blue","blue")
    abcc=(as.integer((input$plot_hover$x-  min(d()))*10  / (max(d())-min(d()) ))+1)
    color[[abcc]]="orange"
    return(color)
    #print(color)
  })

  output$plot <- renderPlot({
    dist <- input$dist
    n <- input$numb
    minv=min(d())
    maxv=max(d())
    hist(d(),breaks=seq(minv,maxv,l=11),main = paste(faa(),n, " Random Variables", sep = ""),col = dkk(), border = "white")
  })
}

# Run the application 
shinyApp(ui = ui, server = server)


#The app can be run on cloud at https://sangamkotalwar.shinyapps.io/Assignmnet_2Final/
```
**The app can be run on cloud [here!](https://sangamkotalwar.shinyapps.io/Assignmnet_2Final/)**
\newpage

#Assignment C

##Q.1] 1. Perform the following steps and comment on the observation.

###Step I. Generate one U(-100,100) random number. Call it m
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
m <- floor(runif(1, -100, 100))
print(paste0("m = ", m))
```
\hfill\break
###Step II. Generate one U(10,50) random number. Call it s.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
s <- floor(runif(1, 10, 50))
print(paste0("s = ", s))
```
\hfill\break
###Step III. Generate one U(10,25) random number. Call it n.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
n <- floor(runif(1, 10, 25))
print(paste0("n = ", n))
```
\hfill\break

###Step IV. Generate 1000 N(m,s) random numbers. Call this the population.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
population <- rnorm(1000, m, s)
print(head(population))
```
\hfill\break

###Step V. Sample n numbers without replacement from the population.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
set.seed(1)
sample <- sample(population, n, replace = FALSE)
print(paste("Head of sample: ", sample))
```
\hfill\break

###Step VI. Construct 90%, 95%, and 99% confidence intervals for the population mean.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(BSDA)
ninty <- z.test(population, sigma.x = s, conf.level = 0.90)
print("For 90% interval: ")
print(ninty)

nintyFive <- z.test(population, sigma.x = s, conf.level = 0.95)
print("For 95% interval: ")
print(nintyFive)

nintyNine <- z.test(population, sigma.x = s, conf.level = 0.99)
print("For 99% interval: ")
print(nintyNine)
```
\hfill\break

###Step VII. Construct 90%, 95%, and 99% confidence intervals for the population variance.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(EnvStats)
df <- length(population) - 1
print(paste("degree of freedom = ", df))
varPopulation <- var(population)
print(paste0("Population variance = ", varPopulation))

print("For 90% interval: ")
print(varTest(population, conf.level = 0.9, sigma.squared = s))

print("For 95% interval: ")
# lowerNintyFive = varPopulation * df / qchisq(0.05/2, df, lower.tail = FALSE)
# upperNintyFive = varPopulation * df / qchisq(1 - 0.05/2, df, lower.tail = FALSE)
# print(c(lower = lowerNintyFive, variance = varPopulation, upper = upperNintyFive))
print(varTest(population, conf.level = 0.95, sigma.squared = s))

print("For 99% interval: ")
print(varTest(population, conf.level = 0.99, sigma.squared = s))
```
\hfill\break

###Step VIII. Repeat steps V & VI 100/500/1000 times and count the number of times (and percentage) that the population mean is captured by the confidence interval.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
draws = matrix(rnorm(floor(n) * 100, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/100 * 100
paste("For n=100: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 500, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/500 * 100
paste("For n=500: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 1000, m, sqrt(s)), n)
get.conf.int = function(x) t.test(x)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= m & conf.int[2, ] >= m)
percentage = count/1000 * 100
paste("For n=1000: Count = ", count, ", Percentage = ", percentage, "%")
```
\hfill\break

###Step IX. Repeat steps V & VII 100/500/1000 times and count the number of times (and percentage) that the population variance is captured by the confidence interval.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
draws = matrix(rnorm(floor(n) * 100, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= s & conf.int[2, ] >= s)
percentage = count/100 * 100
paste("For n=100: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 500, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= s & conf.int[2, ] >= s)
percentage = count/500 * 100
paste("For n=500: Count = ", count, ", Percentage = ", percentage, "%")

draws = matrix(rnorm(floor(n) * 1000, m, sqrt(s)), n)
get.conf.int = function(x) varTest(x, conf.level = 0.95, sigma.squared = s)$conf.int
conf.int = apply(draws, 2, get.conf.int)
count = sum(conf.int[1, ] <= s & conf.int[2, ] >= s)
percentage = count/1000 * 100
paste("For n=1000: Count = ", count, ", Percentage = ", percentage, "%")
```
\hfill\break

##Q.2] In a filament cut test, a razor blade was tested six different times with ultimate forces corresponding to 8.5, 13.9, 7.4, 10.3, 15.7, 4.0.
 
###a] find 95% confidence interval on mean using standard t-distribution
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
forces <- c(8.5, 13.9, 7.4, 10.3, 15.7, 4.0)
forcesNintyFive <- t.test(forces, sigma.x = sd(forces), conf.level = 0.95)
print("For 95% interval: ")
print(forcesNintyFive)
```
\hfill\break
###b] Find a 95% confidence interval on the mean using Efron's percentile method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(boot)
fboot <- function(x, i) mean(x[i])
bs <- boot(forces, fboot, R=1000)
efronForcesPerc <- boot.ci(bs, conf=0.95, type="perc")
print(efronForcesPerc)
```
\hfill\break

###c] Find a 95% confidence interval on the mean using the BCa method and the ABC method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(boot)
fbootBCa <- function(x, i) mean(x[i])
bsBCa <- boot(forces, fbootBCa, R=1000)
efronForcesBCa <- boot.ci(bsBCa, conf=0.95, type="bca")
print("BCa test")
print(efronForcesBCa)

print("ABC test")
fabc <- function(x, w) w%*%x
efronForcesABC <- abc.ci(forces, fabc, conf =0.95)
print(efronForcesABC)
```
\hfill\break

###d] Find a 95% confidence interval on the mean using the percentile-t method.
```{r echo=F, message=FALSE, warning=FALSE, error=FALSE}
library(resample)
bootC <- bootstrap(forces, mean, seed = 0)
percentileForces <- CI.percentile(bootC, probs = c(0.025, 0.975))
# percentileForces <- CI.percentile(bootC, probs = c(0.95))
print("percentile-t  test:")
print(percentileForces)
```
\hfill\break

#Assignment D

##Q.1]
```{r s, echo=FALSE}
sample = c(4.94, 5.06, 4.53, 5.07, 4.99, 5.16, 4.38, 4.43, 4.93, 4.72, 4.92, 4.96)
```

###(a) Estimate an Efron percentile bootstrap 90% confidence interval on the mean aflatoxin residue. 
Use B = 1000 resamples
```{r a, echo=FALSE}
library(boot)
sample.mean<-function(data,x){
    mean(data[x])
}

boot_obj = boot(sample, sample.mean, R=1000)
boot.ci(boot_obj, conf = 0.90, type="perc")
```

###(b) Compare the alfatoxin level found with the industry average value of 5.7 ppm
Is the upper confidence limit less than 5.7 ppb, or is it equal or above? What does this imply about a hypothesis test of H0 : mu >= 5.7 ppb versus H1 : mu < 5.7 ppb at the alpha = 0.05 significance level?

```{r b, echo=FALSE}
boot.ci(boot_obj, conf = 0.95, type="perc")
```


###(c) Find the P-value for the test in (b)

```{r c, echo=FALSE}
z = (mean(sample) - 5.7)/(sd(sample)/sqrt(length(sample)))
2*pnorm(-abs(z))
```

**Result:**

- Since P-value < 0.05, so reject null hypothesis.



##Q.2]

###(a) Find the observed Recall R, Precision P, figure of merit F2.

```{r 2a, echo=FALSE}
recall = 123/(123+27)
precision = 123/(123+625)
figureOfMerit = (precision*recall)/(0.8*precision + 0.2*recall)
print('Precison:')
print(precision)
print('Recall:')
print(recall)
print('F-score:')
print(figureOfMerit)
```

###(b) Resample the 2 × 2 contingency table B = 1000 times. (Hint: Use the multinomialdistribution and rmultinom() in R.)

->
```{r contingency resample}
Relevant_true <- c(123,27)
Irrelevant_true <- c(625,6703)
total_relevant_doc <- 150
total_irrelevant_doc <- 7328
Relevant_true <- rmultinom(1000, total_relevant_doc, Relevant_true)
Irrelevant_true <- rmultinom(1000, total_irrelevant_doc, Irrelevant_true)
```

\hfill\break  

###(c) Find 90% and 95% confidence intervals for the true F2 for the complete database using Efron's percentile method.

->
The 90% confidence interval for F2 of database is: 
```{r contingency CI 90%, echo=FALSE}
f2.boot <- c()
for(i in 1:1000)
  {
    p = Relevant_true[1,i]/(Relevant_true[1,i] + Irrelevant_true[1,i])
    r = Relevant_true[1,i]/150
    f_merit = p*r/(0.8*p + 0.2*r)
    f2.boot <- c(f2.boot, f_merit)
}
quantile(f2.boot,probs = c(0.05,0.95))
```