---
title: "Retail Forecasting"
author: "Prajwal Brijesh Ainapur"
date: "9/23/2018"
output: html_document
---
# CRISP - DM for sales in the company

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = ("~/Assignments/Text"))
```

# Reading the Dataset

```{r input, echo=FALSE}
data <- read.csv("file.txt", header = TRUE, sep = ",")
head(data)
```

# Data Pre-Processing

Step 1: Checking for any Missing Data and Statistical Summary

```{r prepro, echo = FALSE, warning=FALSE, message=FALSE}
library(Hmisc)
library(dplyr)
library(ggpubr)
library(e1071) 
details <- describe(data) 
details
```
Results: 
1: No missing values were found.
2: The highest and lowest values of each parameter in the dataset indicates that those five are almost closer to each other i.e. no outliers.


Histogram Plot for the sales of men's clothing
```{r show2, echo = FALSE, message = FALSE, warning = FALSE}
gghistogram(data$men)
```
Histogram Plot for the sales of women's clothing
```{r show1, echo = FALSE, message = FALSE, warning = FALSE}
gghistogram(data$women)
```


=> Testing for normality using Shapiro-Wilk Test


```{r normality, echo = FALSE, message = FALSE, warning = FALSE}
shapiro.test(data$men)
shapiro.test(data$women)
```
Result: 
1: The p-value for both the sales data is not normally distributed.
2: Data Normalization is required.

ii -> Calculating the trend and removing it for men's sales

```{r seasonality, echo = FALSE, message = FALSE, warning = FALSE}
library(forecast)
trend <- ma(data$men, order = 2, centre = TRUE)
without_trend <- data$men - trend

plot(data$men,type = "l")
trend1 <- ma(without_trend, order = 2, centre = TRUE)
plot(trend1)
without_trend1 <- without_trend - trend1
plot(without_trend1)
```

ii -> Normalizing the data using skewness reduction techniques

```{r skewness, echo = FALSE, message = FALSE, warning = FALSE}
library(forecast)
cat("Skewness of the original data of the women's sales is:",skewness(data$women))
normal_women <- (data$women)^(1/3)
cat("SKewness of the normalized data for the women's sales is:",skewness(normal_women))
```
The skewness of the women's sales data shows a positive skewness of 0.58 and thus is right skewed.
Right skewed data can be reduced by cube root method and as a result of which the skewness was reduced by 10 times.


iii -> Z-Score Standardisation

```{r scaling, echo = FALSE, warning=FALSE, message=FALSE}
scaled_men <- scale(without_trend1, center = TRUE, scale = TRUE)
gghistogram(scaled_men)
ggqqplot(scaled_men, color = "dodgerblue2")

scaled_women <- scale(normal_women, center = TRUE, scale = TRUE)
gghistogram(scaled_women)
ggqqplot(scaled_women, color = "dodgerblue2")
```

Result: The above plot shows that the points now lie closer to the line, after removing the seasonality.

# EDA

```{r show, echo = FALSE, message = FALSE, warning = FALSE}

ggdensity(scaled_men, main = "Density Plot of men sales", xlab = "sales", color = "gray", fill = "dimgray")

ggdensity(scaled_women, main = "Density Plot of women sales", xlab = "sales", color = "gray", fill = "dimgray")

```
=> Testing for normality
```{r check, echo = FALSE, message = FALSE, warning = FALSE}
print(shapiro.test(scaled_men))
print(shapiro.test(scaled_women))
```
Result: The data has a considerably good p-value and thus cleares the test for normality

```{r corr, echo = FALSE, message = FALSE, warning = FALSE}
new_df <- data.frame(data$date, scaled_men, scaled_women)
lat_df <- na.omit(new_df)

cat("Correlation coeffficient between the men's and women's sales is",cor(lat_df$scaled_men, lat_df$scaled_women))
print("Thus the sales of both the genders is closely related.")
```